---
title: "UCM Models"
author: "Jacopo Nicosia"
date: "7/25/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE,message=FALSE, echo = FALSE}

# Libraries & data import

library("xts")
library('Metrics')
library('tseries')
library("dplyr")
library("xts")
library("forecast")
library("urca")
library("lmtest")
library("KFAS")
library("fastDummies")

setwd("C:\\Users\\jacop\\Desktop\\Time Series\\Progetto Jaco")
dt <- read.csv('Dati_orari.csv')

dt$date <- as.Date(as.character(dt$date), format = "%m/%d/%Y")

dt_train <- dt[dt$date <= '2020-05-30',]
dt_val <- dt[dt$date > '2020-05-30',]

```

# Data Import

Train-validation set split.

```{r,warning=FALSE,message=FALSE}

y <- xts(dt_train$h14/10000, as.Date(as.character(dt_train$date), format = "%Y-%m-%d"))
y_val <- xts(dt_val$h14/10000, as.Date(as.character(dt_val$date), format = "%Y-%m-%d"))

```


# Model 1

The first analyzed model is a simple LLT. Its purpose is mainly to set a benchmark to compare future models.

```{r,warning=FALSE,message=FALSE}

mod1 <- SSModel(y ~ SSMtrend(2, list(NA, NA)), H = NA)

vr <- var(y)

fit <- fitSSM(mod1, inits = log(c(vr/10,vr/100,vr/2)))

fit$optim.out

fit$model$Q

```
# Prediction & Evaluation

```{r}

pred1 <- predict(fit$model, n.ahead = 93, interval = "prediction")

mae1 <- mean(abs(as.vector(y_val) - pred1))
mae1

```
As expected the Mean Absolute Error is significantly high. Obviously the model is too simple to explain the selected time series.

# Results Evaluation
```{r,warning=FALSE,message=FALSE}

kfs1<-KFS(fit$model, 
          filtering = c('state','signal'),
          smoothing = c("state", "signal","disturbance"))

ares <- rstandard(kfs1,"state")
plot(ares[,"level"])

```


```{r , fig.width=10, fig.height=6, fig.align='center', echo=FALSE }

plot(y)
lines(xts(kfs1$att[,"level"], time(y)), col = "red")

plot(xts(kfs1$att[,"slope"], time(y)), col = "blue")

```
Thanks to the Kalman Filter is it possible to observe that the time series variability is strongly dependent on the level variability, while the slope stays pretty mich constant.

# Model 2

The following model is much more complex since it includes two types of seasonal components.
- A weekly seasonality, which it is modelled through dummy variables
- A semi-annual seasonality modelled through trigonometric functions (it has been noticed that a small number of harmonics return models with lower MAE)

```{r}

mod2 <- SSModel(as.vector(y) ~ SSMtrend(2, list(NA, NA))+
                      SSMseasonal(7, NA, "dummy") + 
                      SSMseasonal(365, NA, "trig", harmonics = 1:5),
                     H = NA)

vr2 <- var(as.vector(y))

mod2$P1inf <- mod2$P1inf * 0 # Valori di inizializzazione
mod2$a1[1] <- mean(y)
diag(mod2$P1) <- vr2


pars_2 <- numeric(5)

pars_2[1] <- log(vr2/10)       # Level
pars_2[2] <- log(vr2/100)       # Slope
pars_2[3] <- log(vr2/1000)      # Seasonal dummy
pars_2[4] <- log(vr2/1000)      # Seasonal trig
pars_2[5] <- log(vr2/5)       # Obs error

  # Funzione di update
updt_2 <- function(pars, model){
    # Level
    model$Q[1, 1, 1] <- exp(pars[1])
    # Slope
    model$Q[2, 2, 1] <- exp(pars[2])
    # Seasonal Dummy
    model$Q[3, 3, 1] <- exp(pars[3])
    # Seasonal Sinusoids (number of harmonics*2 + 3)
    diag(model$Q[4:13, 4:13,1]) <- exp(pars[4])
    # Observation 
    model$H[1, 1, 1] <- exp(pars[5])
    model
}


```



```{r,warning=FALSE,message=FALSE}

fit2 <-  fitSSM(mod2, pars_2, updt_2)
fit2$optim.out$convergence # if 0, the model has reached convergence

```


```{r,warning=FALSE,message=FALSE}

pred2 <- predict(fit2$model, n.ahead = 93)

```

In the following chunk the different components of the model are visualized for evaluation purposes.

```{r, fig.width=10, fig.height=6, fig.align='center', echo=FALSE }

kfs1 <- KFS(fit2$model, smoothing = c("state", "signal", "disturbance"))


level <- kfs1$alphahat[, "level"]
trig <- kfs1$alphahat[, "sea_trig1"]

plot(as.vector(y), type = "l", 
     ylim = c(min(as.vector(y), kfs1$alphahat[, "level"], kfs1$alphahat[, "slope"]), 
              max(as.vector(y), kfs1$alphahat[, "level"], kfs1$alphahat[, "slope"])))
lines(kfs1$alphahat[, "level"], col = "blue")
lines(kfs1$alphahat[, "slope"], col = "red", lwd = 2)
lines(kfs1$alphahat[, "sea_dummy1"] + level + trig, col = "green", lwd = 2)
lines(kfs1$alphahat[, "sea_trig1"] + level , col = "purple", lwd = 2)

```

It seems that the trigonometric component is much more adaptive then desired. It could be a solution to reduce even more the number of harmonics.

```{r, fig.width=10, fig.height=6, fig.align='center', echo=FALSE }

plot(y_val)
lines(xts(pred2, time(y_val)), col = "red", lwd = 1)

```


```{r,warning=FALSE}

mae_ucm2 <- mean(abs(as.vector(y_val) - pred2))
mae_ucm2

```

# Model 3

In this model the number of harmonics is reduced even further.

```{r,warning=FALSE,message=FALSE}

mod3 <- SSModel(as.vector(y) ~ SSMtrend(2, list(NA, NA))+
                      SSMseasonal(7, NA, "dummy") + 
                      SSMseasonal(365, 0, "trig", harmonics = 1:3),
                     H = NA)

vr3 <- var(as.vector(y))

mod3$P1inf <- mod3$P1inf * 0
mod3$a1[1] <- mean(y)
diag(mod3$P1) <- vr3


```


```{r,warning=FALSE}

fit3 <-  fitSSM(mod3, pars_2)
fit3$optim.out$convergence

```

```{r,warning=FALSE,message=FALSE}

pred3 <- predict(fit3$model, n.ahead = 93)

```

```{r,warning=FALSE}

mae_ucm2 <- mean(abs(as.vector(y_val) - pred3))
mae_ucm2

```
The improvement obtained by reducing the number of harmonics even more is significant.
The previous MAE was $60.94352$ against a new one of $47.29854$.
For this reason the model is trained for each hourly time series.

# Hourly based Final Models
```{r,warning=FALSE,message=FALSE}

modelli <- list()

for (i in 2:25) {
  
  y <- xts(dt_train[[i]]/10000, as.Date(as.character(dt_train$date), format = "%Y-%m-%d"))
  vr <- var(as.vector(y))
  
  mod0 <- SSModel(as.vector(y) ~ SSMtrend(2, list(NA, NA))+
                      SSMseasonal(7, NA, "dummy") + 
                      SSMseasonal(365, 0, "trig", harmonics = 1:3),
                     H = NA)

  mod0$P1inf <- mod0$P1inf * 0 
  mod0$a1[1] <- mean(y)
  diag(mod0$P1) <- vr
  
  pars_0 <- numeric(5)
  
  pars_0[1] <- log(vr/100)       # Level
  pars_0[2] <- log(vr/1000)       # Slope
  pars_0[3] <- log(vr/1000)      # Seasonal dummy
  pars_0[4] <- log(vr/1000)      # Seasonal trig
  pars_0[5] <- log(vr/5)       # Obs error 

  fit0 <-  fitSSM(mod0, pars_0)
  
  modelli[[i - 1]] <- fit0
}
  

```

# Final MAE

Likewise to Arima models, to check that everything went smoothly the average of the 24 MAEs is calculated.

```{r,warning=FALSE}

d <- 0

for (i in 1:24){
  
  pred0 <- predict(modelli[[i]]$model, n.ahead = 93)
  
  ts <- xts(pred0, time(y_val))
  
  b1 <- xts(dt_val[i+1]/10000, as.Date(as.character(dt_val$date)))
  
  c <- mae(ts,b1)
  
  d <- sum(d,c)
}

FinalMAE <- (d/24)*10000

FinalMAE

```

# Final Training
```{r,warning=FALSE,message=FALSE}

modelli <- list()

for (i in 2:25) {
  
  y <- xts(dt[[i]]/10000, as.Date(as.character(dt$date), format = "%Y-%m-%d"))
  vr <- var(as.vector(y))
  
  mod0 <- SSModel(as.vector(y) ~ SSMtrend(2, list(NA, NA))+
                      SSMseasonal(7, NA, "dummy") + 
                      SSMseasonal(365, 0, "trig", harmonics = 1:3),
                     H = NA)

  mod0$P1inf <- mod0$P1inf * 0 ##### Parametri importanti, capire perchè!
  mod0$a1[1] <- mean(y)
  diag(mod0$P1) <- vr
  
  pars_1 <- numeric(5)
  
  pars_1[1] <- log(vr/100)       # Level
  pars_1[2] <- log(vr/1000)       # Slope
  pars_1[3] <- log(vr/1000)      # Seasonal dummy
  pars_1[4] <- log(vr/1000)      # Seasonal trig
  pars_1[5] <- log(vr/5)       # Obs error 
  
  fit0 <-  fitSSM(mod0, pars_1)
  
  modelli[[i - 1]] <- fit0
}
  

```

# Prediction Dataframe Preparation

The following steps are the same already performed for the Arima Model.

```{r,warning=FALSE,message=FALSE}

days <- seq(from=as.Date('2020-09-01'), to=as.Date("2020-10-31"),by='days' )

days1 <- data.frame(days)

ore_iter <- rep(c(1:24),times=61)

k <- 0

for (i in seq_along(days)){
  k <- append(k,rep(days[i], times=24))
}

df1 <- data.frame(as.Date(k[-1]))
df1 <- cbind(df1,ore_iter)
colnames(df1) <- c('date','ora')

```

# Storage
```{r,warning=FALSE,message=FALSE}

for (i in 1:24){
  
  pred0 <- predict(modelli[[i]]$model, n.ahead =61)
  
  UCM <- xts(pred0 * 10000, as.Date(as.character(unique(days1$days))))
  
  df45 <- data.frame(Y=as.matrix(UCM), date=time(UCM))
  
  ora <- rep((i),times=nrow(df45))
  
  df45 <- cbind(df45,ora)
  
  df1 <- merge(x = df1, y = df45, all = TRUE)
  
  }

df1 <- na.omit(df1)

# df1 to CSV

```


```{r, echo=FALSE}

Dati <- read.csv("C:\\Users\\jacop\\Desktop\\Time Series\\Progetto Jaco\\SDMTSA_861812.csv")
Dati$date <- as.Date(as.character(Dati$date), format = "%m/%d/%Y")

```


```{r, echo=FALSE}
#DFF <- cbind(Dati,df1$fit,ML)
#colnames(DFF) <- c("Data", "Ora", "ARIMA", "UCM", "ML")
#write.csv2(DFF, file = 'SDMTSA_861812_FIN.csv')

```

# Final MAE:

$Arima: 621650.9$

$UCM: 356025.4$

$ML-KNN: 563666.9$


